---
title: "Individual Household Electric Consumption"
author: "Jovita L De Loatch"
date: "August 10, 2014"
output: html_document
---

Description: Measurements of electric power consumption in one household 
with a one-minute sampling rate over a period of almost 4 years. 
The goal is to describe how household energy use varies over a 2-day period in February, 2007.
The following descriptions of the 9 variables in the dataset are taken from the UCI web site:
https://archive.ics.uci.edu/ml/datasets/Individual+household+electric+power+consumption on 8/8/14.

Observation Variables:


* Date: Date in format dd/mm/yyyy
* Time: time in format hh:mm:ss
* Global_active_power: household global minute-averaged active power (in kilowatt)
* Global_reactive_power: household global minute-averaged reactive power (in kilowatt)
* Voltage: minute-averaged voltage (in volt)
* Global_intensity: household global minute-averaged current intensity (in ampere)
* Sub_metering_1: energy sub-metering No. 1 (in watt-hour of active energy). It corresponds to the kitchen, containing mainly a dishwasher, an oven and a microwave (hot plates are not electric but gas powered).
* Sub_metering_2: energy sub-metering No. 2 (in watt-hour of active energy). It corresponds to the laundry room, containing a washing-machine, a tumble-drier, a refrigerator and a light.
* Sub_metering_3: energy sub-metering No. 3 (in watt-hour of active energy). 
     It corresponds to an electric water-heater and an air-conditioner.


```{r}
# 0. Import Data for Analysis.
setwd("/Users/jldeloatch/project1/EDA/ExData_Plotting1/")
household_data <- read.table("household_power_consumption.txt", header = TRUE, sep = ";", as.is = TRUE, 
                na.strings = "?", colClasses = c("character", "character", "numeric", 
                        "numeric","numeric" ,"numeric", "numeric", "numeric", "numeric"))

household_data_sub <- subset(household_data, Date == "1/2/2007" | Date == "2/2/2007")
attach(household_data_sub) 
summary(Global_active_power)
```

1. Plot 1 - Histogram of household global minute-averaged active power (in kilowatt)

```{r, echo=FALSE}
#plot1.png 
hist(Global_active_power, col = "red", breaks = 12,
     main = paste("Global Active Power"),  
     xlab = "Global Active Power (kilowatts)")
```

2. Plot 2 - Time series plot of global minute-averaged active power (in kilowatt)
Merge 'Date' and 'Time'  'Date_Time' column for use of time format

```{r}
#plot2.png
Date_Time <- paste(Date, Time)
Date_Time <- strptime(Date_Time, "%d/%m/%Y %H:%M:%S")

plot(x = Date_Time, y = Global_active_power, type = "l", 
     ylab = "Global Active Power (kilowatts)")

```

3. Plot 3 - Time series plot of energy usage across various sections of the house.
```{r}
#plot3.png
plot(x = Date_Time, y = Sub_metering_1,  type = "l", ylab = "Energy sub metering")
lines(x = Date_Time, y = Sub_metering_2, col = "red")
lines(x = Date_Time, y = Sub_metering_3, col = "blue")
legend("topright", legend = c("Sub_metering_1", "Sub_metering_2", "Sub_metering_3"), 
       lty = 1, col = c("black", "red", "blue"))

```

4. Plot 4 - Combined Time Series Plots
```{r}
#plot4.png
## Split the device into four plotting regions
par(mfcol = c(2, 2))

## 'Global Active Power' vs 'Time' Plot
plot(x = Date_Time, y = Global_active_power, type = "l", 
     ylab = "Global Active Power")

## 'Energy sub metering' vs 'Time' Plot
plot(x = Date_Time, y = Sub_metering_1,  type = "l", ylab = "Energy sub metering")
lines(x = Date_Time, y = Sub_metering_2, col = "red")
lines(x = Date_Time, y = Sub_metering_3, col = "blue")
legend("topright", legend = c("Sub_metering_1", "Sub_metering_2", "Sub_metering_3"), lty = 1, bty = "n", col = c("black", "red", "blue"), cex=.5 )

## 'Voltage' vs 'Time' Plot
plot(x = Date_Time, y = Voltage, type = "l", xlab = "datetime", 
     ylab = "Voltage")

# 'Global_reactive_power' vs 'Time' Plot
plot(x = Date_Time, y = Global_reactive_power, 
     type = "l", xlab = "datetime", ylab = "Global Reactive Power")
```
